---
title: "Analysis Report 1: Forensic Identification using single taxa distribution of skin bacterial communities "
author: "Maggie Chen"
date: "October 20, 2017"
output: github_document
bibliography: references.bib
csl: bioinformatics.csl
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction
Bacteria community constantly interacts with human communities, and it has both positive and negative impact on humans. For instance, microbial community in human gut is a typical example of how human can benefit from bacteria. The microbial community in human's gut support metabolic activities, which will help the provide energy and provide nutrients for human body. The microbial community in gut also help provide immune protection against alien microbes[@mcmurdie2013]. There also a lot of diseases caused by bacteria. For example, periodontist as an oral infection may be caused by bacteria infections, which is even related to systemic diseases, like cardiovascular disease[@guarner2003gut]. Not only bacteria can impact human health, human can also function as a mediator of bacteria trans infection. At different body parts there is a variation of microbial community composition. The human intestinal habitat contains 300 to 500 different bacteria species, which performs different functions. Around 90% of the adult gut bacteria community is is composed of *Firmicute* or *Bacteroidetes* phyla, and the rest of it belongs to *Actinobacteria*, *Proteobacteria*, and several other phyla [@huttenhower2012structure]. However, skin provides a different habitat than gut, thus even in the same individual there is a different microbial community composition. This different is caused by many factos like oxygen level, PH, and immunological factors [@faust2012microbial]. Different from gut, skin is dominated by *Staphylococcus* (Phylum Firmicutes), *Propionibacterium* or *Corynebacterium*(phylum *Actinobacteria*)[@huttenhower2012structure]. Human skin is a good carrier of bacteria, and transfer them onto surfaces through touching. Also, individual's living environment and life style impacts the bacteria community on each person, which creates a significant variations between individuals [@fierer2010forensic]. What's more, skin bacteria communities are relative stable over times [@fierer2008influence]. In addition, bacteria is capable to persist on the touched surface for certain amount of time, even under harsh storage conditions[@fierer2010forensic]. This long time persistence allows re-examination of microbial community available within certain time period. 

Along with the fast development of sequencing techniques, the sequence and investigation on microbial communities has expanded extensively. High-throughtput sequencing, also known as the Next Generation sequencing(NGS) technology, that used non-Sanger-based method to enable a unprecendented DNA sequencing speed. This also allows the novel biological sequencing applications. For example, 454 technology is known to have a key advantage of combing the emulsion PCR and pyrosequencing. For emulsion PCR, the single DNA amplification is a bias-free process, by avoiding the competition by multiple template for a relative limited amount of DNA polymerase. Pyrosequencing, can be performed in parallel and produce a visual signal via computer system [@schuster2007next]. The next generation seuqencing techinique makes the DNA sequencing faster, which generates a significant more data for scientists to proess. Thus, to process the massive amount of data, different methods are developed. One of the example and will be usd in this paper is the Phylogenetic Sequencing (Phyloseq), which is a contingency table of taxonomic and samples, and intergraate with other data types. This can be applied along with the powerful statistical and graphical tool R to enable more diversed data annlysis [@mcmurdie2013].

In addition to the health implication of understand bacteria, another relative new application of understanding microbe community is in the forensic examinations. Recent research has indicated that microbial community can serve as an additional evidence in forensic investigations of suspect identification[@fierer2010forensic]. The combination of the high throughput sequencing technology and the characteristics of bacteria communities on human skins and how they persist on touched surface makes it possible as a forensic identification method. Through comparing and matching the bacteria community composition, the microbial community will serve as an "fingerprint" for the individual identification during forensic investigations [@fierer2010forensic].

The goal for this article, is to explore additional approach of analyze microbial community for forensic identification supposes. My primary question is that instead of matching multiple communities between human skin and object been touched, how single taxa analysis can help with forensic identification. My hypothesis is that in the male community, for the most abundant bacteria taxa will has unique distribution on object, in this case, the keyboards. This unique bacteria distribution will then function as an "fingerprint" for individual identification. In order to investigate on this hypothesis, R program is used for sequence analysis.  
 


# Methods

## Sample origin and sequencing

The samples are taken from three individuals in between age 20 and 35 by Fierer et al., that 25 to 30 keys on three person's personal computer keyboards are swabbed. Fierer et al. also swabbed their skin on the ventral surface of each fingertip's distal joint. As for the sample control, all three individuals are health and had not taken any antibiotics for at least past 6month. Two of the individuals shared the same office space. Then communities on the keys was compared to the communities to the bacteria communities on the owner's fingertip. For the purpose of quantify the degree of association between keyboard and it owner, versus keyboards not touched by the individual, Fierer et al. also sampled other individual keys from other public and private keyboards located on the University of Colorado campus. The sampling tools for both skin and keyboard keys is the autoclaves cotton-tipped swabs, that is premoistened by s sterile solution [@fierer2008influence]. As for the storage of the samples before DNA extraction, all swabs were stored at -80ºC for less than a week [@fierer2010forensic].

Then, the bacterial DNA was extracted from the swabbed samples and the composition of the bacteria community was determined by using the bar-code pyroseqyencing technology. Fierer et al. used MO BIO PowerSoil DNA Isolation kit to extract genomic DNA from the swabs. For each sample, Fierer et al. amplified 16S rRNA genes using specific primers, which was optimized for the phylogenetic analysis. Fierer et al. also used PCR reaction in triplicated reactions with forward and reverse primers. The amplicon yielded from the PCR reaction was visualized on 0.1% agarose gels. Amplicons was then cleaned using the PCR Clean-up kit. The concentration of amplicon DNA were measured using the QUant-iT PicoGreen dsDNA reagent and kit. 454 Life Sciences Genome Sequencer FLX instrument performed by the Environmental Genomics Core Facility at the University of South Carolina for pyrosequencing [@fierer2010forensic]. 


## Computational

I used DADA to analyze and clean the data, then phlyoseq and ggplot to visualize the results. Firstly, I ran a quality control on all twenty of the sample data, which generated a quality profile on each of the samples. Then according to the sequence quality, I trimmed the data if they have more than three expected error, and has a quality lower than two. As a quality control, I created a table to compare the read counts before and after the trim to ensure I did not trim off too many sequences. After the trim, I build an error models for each sample to assess the probability of an unmatched base is an error or not, and a plot to check whether the error model match the data. After it, I also deleted the duplicated sequences. Then as a very important step, I DADA to process the sequence data. DADA is a analysis include the error modle and evaluate whether the unique sequence found is true or false. DADA eliminates the false positive results to provide the real variants and limit spurious sequence, to provide a high resolution amplicon data [@callahan2016].

Then I listed the output of the DADA analysis on a table according to the various of length. As well as a histogram to indicate the distribution of the sequence that is trimmed and denoised, to ensure that all the sequencese are in accetpable length and quality. Using this output, I also removed all the chimeras, which is the cDNA sequence that generated from two transcript that is caused by polymerase stuck and ump to another similar sequence. Using the sequence data, I assigned the taxonomy according to the supplied training set. I also created phylogeny tree using the trimmed and denoised sequence data. Using the dada2 output, I charted a data set, which contains the metadata for each sample, Phylogeny and  taxonomy for each sequence. Then to enable the ggplot function, I combined the taxa and metadata into one table. 


# Results

To answer the primary question that how can a single taxa function as an additional forensic identification evidence. I first exampled the different abundance of phylum found on each individual’s samples. I categorized each individual’s samples into two types, samples from individual skin and keyboards. Throughout all individuals’  keyboards samples, *Actinobacteria* has significant higher abundance across all individuals. The second commonly found Phylum is *Firmicutes* (Figure1). As for the skin samples, only three individuals has skin sample available, for which it is consistent with keyboard samples that the  *Actinobacteria* has a significant high abundance in all three individuals(Figure1). Since *Actinobacteria*, and  *Firmicutes* are two most commonly found bacteria phylum, I further plotted their distributions on each keys on the keyboards. Only three male individuals has all keys sampled, they are male individuals M2, M3, M9. For other individuals, only space bar is sampled. The highest amount of *Actinobacteria* is found on fingertip. M2, M3, M9 has a different distribution pattern of both bacteria on the keyboards(Figure2). *Actinobacteria* is found in a higher abundance across the keyboards and individual. Thus, to further investigate *Actinobacteria*, I plotted the families within phylum *Actinobacteria* according to the abundance found on each of the keys and fingertip. *Propionibacteriaceae* family has a significant higher abundance in all individual sample across all keys and fingertips(Figure2). For M3 and M9, fingertip is the location that has most of *Propionibacteriaceae* found. For M2, all keys and fingertip has relative lower abundance than M3 and M9. All three individuals have different distributions of Propionibacteriaceae on keys(Figure2). Since most of the family identified in samples are *Propionibacteriaceae*, and there are still distinguishable difference in between each individual's bacteria distribution pattern on the keyboard. I further examined the Genus found on three male individuals. The Genus found are concentrated in *Propionibacterium*, and for M3, M9 two male individuals, the fingertip has the highest abundance. For all three individuals there is also a different abundance distributions on the keys of the keyboards(Figure4). Then, the alpha diversity is measured for the sequences identified as Genus *Propionibacterium*. The samples are from both fingertip and the keyboard of the individual. The one of the sample (M2) has a higher diversity in this Genus. And the fingertip in the same individual (M2) has higher diversity than its keyboard. For other two individuals (M3,M9), they have lower diversity in the samples compare to M2. Also, for both M3 and M9, the fingertip and keyboards has a very similar alpha diversity(Figure5).  


```{r load-libraries, message = FALSE}
# Be sure to install these packages before running this script
# They can be installed either with the intall.packages() function
# or with the 'Packages' pane in RStudio

# load general-use packages
library("dplyr")
library("tidyr")
library("knitr")
library("ggplot2")

# this package allows for the easy inclusion of literature citations in our Rmd
# more info here: https://github.com/crsh/citr
# and here:
# http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html
library("citr")

# These are the primary packages well use to clean and analyze the data
# this package needs to be installed from bioconductor -- it's not on CRAN
# see info here: https://benjjneb.github.io/dada2/dada-installation.html
library("dada2")

# This to export a fasta of our final denoised sequence variants
library("seqinr")

# To install this you have to install from GitHub
# See more info here: https://github.com/leffj/mctoolsr
# run this -- install.packages("devtools")
# and then this -- devtools::install_github("leffj/mctoolsr")
library("mctoolsr")

# And this to visualize our results
# it also needs to be installed from bioconductor
library("phyloseq")
```

```{r extract-sample-and-file-names}
# NOTE: Much of the following follows the DADA2 tutorials available here:
# https://benjjneb.github.io/dada2/tutorial.html
# Accessed October 19, 2017

# set the base path for our input data files
path <- "data/raw_data"

# Sort ensures samples are in order
filenames_forward_reads <- sort(list.files(path, pattern = ".fastq"))

# Extract sample names, assuming filenames have format: SAMPLENAME.fastq
sample_names <- sapply(strsplit(filenames_forward_reads, "\\."), `[`, 1)

# Specify the full path to each of the filenames_forward_reads
filenames_forward_reads <- file.path(path, filenames_forward_reads)
```

```{r check-quality-plots, include = FALSE}
# Plots the quality profiles of all twenty samples
plotQualityProfile(filenames_forward_reads[1:115])
```

We can see from the quality profiles that most reads tend to get pretty bad in quality after around 200 bases. 

```{r filter-reads, include = FALSE}
# Place filtered files in filtered/ subdirectory
# note this will fail if the directory doesn't exist
filter_path <- file.path("output", "filtered")
filtered_reads_path <- file.path(filter_path,
                                 paste0(sample_names,
                                        "_filt.fastq.gz"))

# See ?filterAndTrim for details on the parameters
# See here for adjustments for 454 data:
# https://benjjneb.github.io/dada2/
#     faq.html#can-i-use-dada2-with-my-454-or-ion-torrent-data
filtered_output <- filterAndTrim(fwd = filenames_forward_reads,
                                 filt = filtered_reads_path,
                                 maxLen = 300,
                                 maxN = 0, # discard any seqs with Ns
                                 maxEE = 3, # allow w/ up to 3 expected errors
                                 truncQ = 2, # cut off if quality gets this low
                                 rm.phix = TRUE,
                                 compress = TRUE,
                                 multithread = FALSE)
```

```{r filtered-read-counts-table, include = FALSE}
# produce nicely-formatted markdown table of read counts
# before/after trimming
kable(filtered_output,
      col.names = c("Reads In",
                    "Reads Out"))
```

```{r learn-errors, include = FALSE}
# this build error models from each of the samples
errors_forward_reads <- learnErrors(filtered_reads_path,
                                    multithread = FALSE)
```

```{r visualize-errors-with-plots,include = FALSE}
# quick check to see if error models match data
# (black lines match black points) and are generally decresing left to right
plotErrors(errors_forward_reads,
           nominalQ = TRUE)
```

```{r dereplicate-sequences, include = FALSE}
# get rid of any duplicated sequences
dereplicated_forward_reads <- derepFastq(filtered_reads_path,
                                         verbose = TRUE)

# Name the derep-class objects by the sample names
names(dereplicated_forward_reads) <- sample_names
```

```{r run-dada, include = FALSE}
# parameters adjusted based on recommendations for 454 data here:
# https://benjjneb.github.io/dada2/
#     faq.html#can-i-use-dada2-with-my-454-or-ion-torrent-data
dada_forward_reads <- dada(dereplicated_forward_reads,
                           err = errors_forward_reads,
                           HOMOPOLYMER_GAP_PENALTY = -1, # reduce penalty bc 454
                           BAND_SIZE = 32) # performs local alignments bc indels

# check dada results
dada_forward_reads
```

```{r make-sequence-table, include = FALSE}
# produce the 'site by species matrix'
sequence_table <- makeSequenceTable(dada_forward_reads)
```

The output table has `r nrow(sequence_table)` rows (samples) and `r ncol(sequence_table)` columns (sequence variants). Notice how we can embed R code directly in our markdown text.

```{r histogram-of-sequence-lengths, include = FALSE}
# Quick check to look at distribution of trimmed and denoised sequences
hist(nchar(getSequences(sequence_table)),
     main = "Histogram of final sequence variant lengths",
     xlab = "Sequence length in bp")
```

```{r remove-chimeras, include = FALSE}
# Check for and remove chimeras
sequence_table_nochim <- removeBimeraDenovo(sequence_table,
                                            method = "consensus",
                                            multithread = FALSE,
                                            verbose = TRUE)

# What percent of our reads are non-chimeric?
non_chimeric_reads <- round(sum(sequence_table_nochim) / sum(sequence_table),
                            digits = 4) * 100
```

After removing chimeras, we were left with `r non_chimeric_reads`% of our cleaned reads.

```{r table-of-pipeline-read-counts, include = FALSE}
# Build a table showing how many sequences remain at each step of the pipeline
get_n <- function(x) sum(getUniques(x)) # make a quick function
track <- cbind(filtered_output, # already has 2 columns
               sapply(dada_forward_reads, get_n),
               rowSums(sequence_table),
               rowSums(sequence_table_nochim))

# add nice meaningful column names
colnames(track) <- c("Input",
                     "Filtered",
                     "Denoised",
                     "Sequence Table",
                     "Non-chimeric")

# set the proper rownames
rownames(track) <- sample_names

# produce nice markdown table of progress through the pipeline
kable(track)
```

```{r assign-taxonomy, include = FALSE}
# assigns taxonomy to each sequence variant based on a supplied training set
# made up of known sequences
taxa <- assignTaxonomy(sequence_table_nochim,
                       "data/training/rdp_train_set_16.fa.gz",
                       multithread = FALSE,
                       tryRC = TRUE) # also check with seq reverse compliments

# show the results of the taxonomy assignment
unname(taxa)
```

```{r extract-sequences-to-fasta}
# we want to export the cleaned, trimmed, filtered, denoised sequence variants
# so that we can build a phylogeny - we'll build the phylogeny outside of R
# but we need the fasta file to do so. We keep the names of each sequence as the
# sequence itself (which is rather confusing), because that's how DADA2 labels
# it's columns (e.g. 'species')
# function taken from https://github.com/benjjneb/dada2/issues/88
export_taxa_table_and_seqs <- function(sequence_table_nochim,
                                       file_seqtab,
                                       file_seqs) {
  seqtab_t <- as.data.frame(t(sequence_table_nochim)) # transpose to data frame
  seqs <- row.names(seqtab_t) # extract rownames
  row.names(seqtab_t) <- seqs # set rownames to sequences
  outlist <- list(data_loaded = seqtab_t)
  mctoolsr::export_taxa_table(outlist, file_seqtab) # write out an OTU table
  seqs <- as.list(seqs)
  seqinr::write.fasta(seqs, row.names(seqtab_t), file_seqs) # write out fasta
}

# actually run the function, with the names of the files we want it to create
# and where to put them
export_taxa_table_and_seqs(sequence_table_nochim,
                           "output/sequence_variants_table.txt",
                           "output/sequence_variants_seqs.fa")
```


```{r read-in-metadata-and-create-phyloseq}
# Next we want to read in the metadata file so we can add that in too
# This is not a csv file, so we have to use a slightly different syntax
# here the `sep = "\t"` tells the function that the data are tab-delimited
# and the `stringsAsFactors = FALSE` tells it not to assume that things are
# categorical variables
metadata_in <- read.table(paste0("data/metadata/",
                    "fierer_hand_bacteria_SRA_study_ERP022626_SraRunTable.txt"),
                          sep = "\t",
                          header = TRUE,
                          stringsAsFactors = FALSE,
                          row.names = 6) # sets sample IDs to row names

# read in the phylogeny, which was created from the fasta exported above
# in Geneious by aligning the sequences with MAFFT and then building a
# Maximum-Likelihood tree with RAxML
tree_in <- read_tree("output/sequence_variants_MAFFT_FastTree.newick")

# Construct phyloseq object (straightforward from dada2 outputs)
phyloseq_obj <- phyloseq(otu_table(sequence_table_nochim,
                                   taxa_are_rows = FALSE), # sample-spp matrix
                         sample_data(metadata_in), # metadata for each sample
                         tax_table(taxa), # taxonomy for each sequence variant
                         phy_tree(tree_in)) # phylogeny from sequence variants
```



```{r, combine_tables}
# Combine the taxa and metedata into one table
# Enable ggplot function to use both tables information at the same time
melted_obj <- psmelt(phyloseq_obj)
```

```{r, host-phylum-abuandace-barplot}
# Phylum distribution for each individual
# Also, in sepeartion of sample types
  melted_obj %>%
  ggplot(aes(x = host_subject_id,
             y = Abundance,
             fill = Phylum, color = Phylum)) +
  geom_col(position = position_dodge()) +
  facet_grid(sample_type ~.) +
  ggtitle("Abundance of Phylums Found on Individuals")

```

**Figure1**: Abundance of taxa found on each individual and their keyboard samples, colored by Phylum. Separate by sample types, skin as the fingertip and surface as the keyboard. As the Figure presents, the most commonly found Phylum across all the individuals' keyboard samples is *Actinobacteria*, present in orange color. The second commonly found bacteria Phylum is *Firmicutes*. This is consistent with fingertip samples, however, fingertip samples are only available for M2, M3, M9 individuals, who are all males. Other individuals does not have fingertip sample available, including F1 who is a female individual; for L1, L3, R1, U1, U2, U3 the sex of the individual is not applicable.


```{r, Actinobacteria-Firmicutes_key_identification-barplot}
# Filter the most commonly found species, Actinobacteria and Firmicutes
# and examine the sample source of where Actinobacteria and Firmicutes was found
melted_obj %>%
  filter(Phylum %in% c("Actinobacteria", "Firmicutes")) %>%
  ggplot(aes(x = host_subject_id,
             y = Abundance,
             fill = sample_source,
             color = sample_source)) +
  geom_col(position = position_dodge()) +
  facet_grid(Phylum ~.) +
  ggtitle("Abundance of Actinobacteria and Firmicutes Found on Keys")

```

**Figure2**: Abundance of two most commonly found bacteria phylum on each individuals, separated by Phylum *Actinobacteria* and *Firmicutes*, colored by sample source. The sample source include each of the keys on the keyboard and individual fingertip samples. As an result, only male individuals M2, M3, M9 has all keys sampled. For other individuals, only space bar is sampled. The highest amount of Actinobacteria is found on fingertip. M2, M3, M9 has a different distribution pattern of both bacteria on the keyboards. 



```{r, male-sample_type_Family-barplot}
# To examine for male indivudals phylums of Actinobacteria
# then look for the Family distribution on each keys
# for each of the male samples
melted_obj %>%
  filter(host_subject_id %in% c("M2", "M3", "M9")) %>%
  filter(Phylum %in% c("Actinobacteria")) %>%
  ggplot(aes(x = Family,
             y = Abundance,
             color = sample_source,
             fill = sample_source)) +
    geom_col(position = position_dodge()) +
  facet_grid(host_subject_id ~.) +
  theme(axis.text.x = element_text(angle = 90,
                                    hjust = 1)) +
  ggtitle("Abundance of Families found on Different Keys
          for M2, M3, M9")

```

**Figure3**: Abundance of Families within phylum Actinobacteria on M2, M3, M9 individuals, separate by individuals, and colored by sample source. The sample source include each of the keys on the keyboard and individual fingertip samples. *Propionibacteriaceae* family has a significant higher abundance in all individual sample across all keys and fingertips. For M3 and M9, fingertip is the location that has most of Propionibacteriaceae found. For M2, all keys and fingertip has relative lower abundance than M3 and M9. All three individuals have different distributions of *Propionibacteriaceae* on keys. Only M2 has a small amount of *Corynrbacteriaceae* has identified in some of the keys and most of the fingertip. 


```{r, Genus-in-Propionibacteriaceae_Family-barplot}
# Filter the Family Propionibacteriaceae, which is most common
# Further look for Genus distribution in 3 male samples
# on each of the keys examed
melted_obj %>%
  filter(host_subject_id %in% c("M2", "M3", "M9")) %>%
  filter(Family == "Propionibacteriaceae") %>%
  ggplot(aes(x = Genus,
             y = Abundance,
             color = sample_source,
             fill = sample_source)) +
    geom_col(position = position_dodge()) +
    facet_grid(host_subject_id ~.) +
  theme(axis.text.x = element_text(angle = 90,
                                    hjust = 1)) +
  ggtitle("Abundance of Genus in Propionibacteriaceae
          family on M2, M3, M9")

```

**Figure4**: Genus abundance in *Propionibacteriaceae* family on M2, M3, M9. For all three male individuals M2, M3, M9, the most common Genus identified on both their keyboards and fingertip is *Propionibacterium*. For individual M3 and M9, the most abundance on found on Fingertip. For each of the individual, there is a different distribution of amount of Genus *Propionibacterium* found on each of the keys.

```{r, subset-Genus}
# use subset data to generate a selected dataset
# for the alpha diversity check
# create a subset for only Genus Propionibacterium
subset_genus <- subset_taxa(phyloseq_obj, Genus == "Propionibacterium")

```

```{r, subset-Male}
# use subset data to generate a selected dataset
# for the alpha diversity check
# make another subset that only contains information
# of Genus Propionibacterium for M2,3,9
subset2_male <- subset_samples(subset_genus,
                               host_subject_id %in% c("M2", "M3", "M9"))
```

```{r, jitterplot-sample-diversity}
# Use created subset data to plot for the
# alpha diversity metirc on different keys
plot_richness(subset2_male,
              x = "sample_type",
              measures = c("Shannon"),
              color = "sample_source") +
  xlab("Sample origin") +
  geom_jitter(width = 0.2) +
  facet_grid(host_subject_id ~.) +
  theme_bw() +
  ggtitle("Alpha Diversity of Genus Propionibacterium on M2, M3, M9 Samples")
```

**Figure5**: The male sample alpha diversity of sequences identified as Genus *Propionibacterium*, colored by sample source. This figure presents the alpha diversity on sample from M2, M3, and M9, which are all male individuals. The alpha diversity is measured for the sequences identified as Genus *Propionibacterium*. The samples are from both fingertip and the keyboard of the individual. The M2 has a higher diversity in this Genus. And the fingertip in M2 has higher diversity than its keyboard. For M3 and M9, they have lower diversity in the samples compare to M2. Also, for both M3 and M9, the fingertip and keyboards has a very similar diversities. 





# Discussion

As the result of the quality check, all the sequences are in good quality overall, specifically, all sequences remains in good quality until it pass the two hundred base. As the result of trimming, due to the sequences are generally in good qualities, most part of each each sequence has survivied the trimm process. As well as the dada2 process. Thus, all the sequences used in the analysis are vaild, which mean the observation and results I have reached is hgihly reliable. 

*Firmicutes* and *Actinobacteria* are two phylums commonly found on human skin, especially *Propionibacterium* and *Corynebacterium* within phylum *Actinobacteria* [@huttenhower2012structure]. As an result, the most commonly found bacteria I observed across all samples are *Firmicutes* and *Actinobacteria*. Also, the one Family and Genus the bacteria has commonly found are *Propionibacterium* and *Corynebacterium*. What's more, the bacteria *Actinobacteria* and *Firmicutes* are found both on fingertip and the keyboards, which support the idea that human skin will pass bacteria onto objects through touching [@fierer2008influence]. Thus, the observation is consistant with previous research results. Furthermore, due to the fact that *Actinobacteria* has a significant higher abundance, and it is found across all individuals, including control individuals, who only has space bar sampled. Thus, as a the primary hyposthsis that whether a single taxa can function as a forensic idnetifier, *Actinobacteria* with high abundance will provide a sufficient amount of samples, which reduces the possilitiy of flase positive results. 

The most commonly found Genus *Propionibacterium* is a gram-positive and anaerobic bacteria that is primarily facultative parasites and commensals of human. It lives around different areas of skin, including sweat glands, and sebaceous glands. *Propionibacterium* is a virtually ubiquitous, and cause no problem for most of human population, but *Propionibacteria* have been invloveld in acne an other skin conditions [@bojar2004acne]. According to this, not even mention the share of keyboards might cause the transinfection of skin conditions through bacteria, for personal health purposes, clean keyboards or hand on regular basis will be necessary to limit the bacteria growth and cross inffections of skin conditions. 

The observation made in the analysis has confirmed on the hypothesis that one taxa has the ability to function as forensic identification evident, by using the abundance distribution pattern of one taxa on the object(Keyboard). In this analysis, I have observed the distribution variation between three individals at three levels, Phylum(Figure2), Family(Figure3), and Genus(Figure4), which serves as a strong support that one taxa's distribution will present individual difference. Furthermore, the taxa being used in this distribution pattern analysis cannot be the taxa has low abundance on the object. *Actinobacteria* has found with a significant high abundance across all individuals in comparison. Also, in all sample bacteria finding are concentrated in one Family and one Genus, in which there is a difference in distribution pattern. This variant distribution pattern might be caused by the different personal habit of using keyboards, like finger perferences in typing. This additional approach has a vital due to sometimes, individuals who works in the same environment, will be very likely to have similar microbial community composition, but by including this single bacteria taxa distribution pattern, it can provide a in depth analysis for forensic identification. 

For three samples in my interest two of the individuals have relatively high overall abundance of bacteria found on keyboard, and both of them has the highest taxa abundance found on the fingertip, the only difference is the distribution pattern. Both of them also as a relative low alpha diversity of the Genus found on the keyboard. According to the sampling method of the initial experiment, there are two sampling individuals worked in the same office areas, this can  be the cause of the similarities between their bacterial community. For the other individual (M2), who might work in a different office, thus, only he has an small amount of *Corynrbacteriaceae* found on his fingertip and some of the keys. The higher alpha diversity in the M2 in indicate the possibility of have different strains or different species that is hard for sequence to detect, which might be caused by the working environment he has contacted with.

There are several things can be done to further investigate this single taxa distribution patter identification method. One is that, there are only three individuals has all keys on the keyboard samples, repeating the experiment with more individuals will sever as a better support of the conclusion. Another one is that all three samples are from male individuals, by including females samples can create another layer of the experiment that whether sex will impact the microbial community composition and whether there is a distribution pattern variation between female individuals. 

# Sources Cited


